---
title: "peaks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{peaks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
# library(mcATAC)
devtools::load_all("/home/aviezerl/src/mcATAC/", export_all = FALSE)
library(ggplot2)
ggplot2::theme_set(ggplot2::theme_classic())
```

## Download fragments data

We will start by downloading the fragments data (and it's index) from 10x:

```{r download-data}
if (!dir.exists("pbmc_data") || !file.exists("pbmc_data/fragments.tsv.gz") || !file.exists("pbmc_data/fragments.tsv.gz.tbi")){
  download_pbmc_example_data(fragments = TRUE)
}
```

This would download the fragments file and its index from:

[PBMC from a healthy donor - granulocytes removed through cell sorting (10k)](https://support.10xgenomics.com/single-cell-multiome-atac-gex/datasets/1.0.0/pbmc_granulocyte_sorted_10k?)

Note that the 'fragments' file is large - ~2GB, so this step might take some time. 

## Import raw reads into a ScCounts format

`mcATAC` allows you to load the raw counts data into an `ScCounts` object. A `ScCounts` holds the raw reads as a set of sparse matrices where the rows are genomic coordinates and the columns are cell barcodes. The number of sparse matrices ("genomic bins") is controlled by the `bin_size` parameter. The matrices are stored in a 'gzip' compressed form under a directory called 'data', together with a 'yaml' file containing the some metadata such as the genomic bins coordinates and the cell barcodes.

We can now import the raw reads into a `ScCounts` object that we are going to store at "pbmc_data/reads". Note that this might take ~4 minutes on a machine with 96 cores. 
  
```{r import-scounts, eval = FALSE}
barcodes <- tgutil::fread("pbmc_data/barcodes.tsv", header = FALSE)[, 1]
write_sc_counts_from_fragments("pbmc_data/fragments.tsv.gz", "pbmc_data/reads", cell_names = barcodes, genome = "hg38", overwrite = TRUE)
```

## Load ScCounts object

```{r load-scounts}
sc_counts <- scc_read("pbmc_data/reads")
sc_counts
```

## Project RNA metacells

We can now project metacells derived from the RNA data (or any other way) into an `McCounts` object which would hold the counts per metacell: 

```{r project-mc}
data(cell_to_metacell_pbmc_example)
mc_counts <- scc_project_on_mc(sc_counts, cell_to_metacell_pbmc_example)
mc_counts
```

## Save McCounts object

```{r save-mc-counts}
mcc_write(mc_counts, "pbmc_data/mc_counts", overwrite = TRUE)
```

## Create 'misha' tracks for each metacell 

We can now create a 'misha' track for each metacell:

```{r misha-tracks}
mcc_to_tracks(mc_counts, "pbmc_mc", overwrite = TRUE)
```

## Identify peaks

We can then use the tracks we created to identify peaks on the metacell level. We think it is more robust than looking only at the marginal of all the data (as the cell ranger does), or than trying to identify peaks on the single cell level which is very sparse. 

```{r identify-peaks}
# identify peaks on the metacell level. In the meantime - a placeholder using the ScATACPeaks peaks
peaks <- atac_sc@peaks
```

## Create a McATACPeaks object from McCounts object 

After we have indentified the peaks, we can create a `McATACPeaks` object from the `McCounts` object that we have created.

```{r mc-atac}
atac_mc <- mcc_to_mcatac(mc_counts, peaks)
atac_mc
```



